A Regularized Time Varying Vector Error Correction Model

### 2016-03-29 articles-conferences/elsevier-2016/src/background.tex
avance martes

VECM is a special case of VAR model and both describe the joint behaviour of a
set of variables.

VAR($p$) model is a general framework to describe the behaviour of a set of $l$
endogenous variables as a linear combination of their last $p$ values. These
$l$ variables at time $t$ are represented by the vector $\mathbf{y}_t$ as
follows:

\begin{equation}
\label{eq:variables}
\mathbf{y}_t = 
\begin{bmatrix} y_{1,t} &
y_{2,t} &
\dots &
y_{l,t}
\end{bmatrix}^\top \, ,
\end{equation}
\noindent where $y_{j,t}$ corresponds to the time series $j$ evaluated at
time $t$.

The VAR(p) model describes the behaviour of a dependent variable in terms
of its own lagged values and the lags of the others variables in the
system. The model with $p$ lags is formulated as the following:

\begin{equation}
\label{eq:var}
 \mathbf{y}_t = \phi_1 \mathbf{y}_{t-1}  + \dots +   \phi_p\mathbf{y}_{t-p}
+ \mathbf{c} + \mathbf{\epsilon}_t \, ,
\end{equation}

\noindent where $t=p+1, \dots N$,  ${\phi_1,\dots,\phi_p}$ are $l \times l$
matrices of real coefficients,
$\mathbf{\epsilon}_{p+1},\dots,\mathbf{\epsilon}_N$ are error terms,
$\mathbf{c}$ is a constant vector and $N$ is the total number of samples.

The VAR matrix form of equation~(\ref{eq:var}) is:

%\begin{equation}\label{eq:vareq}
%\mathbf{B} = \mathbf{A} \mathbf{X} + 
%\mathbf{E} \, , 
%\end{equation}
%
%\noindent where:
%\begin{eqnarray}
%\mathbf{A} &=&
%\begin{bmatrix}
%   \mathbf{y}_{p}     & \dots    & \mathbf{y}_{N-1}\\
%   \mathbf{y}_{p-1}   & \dots    & \mathbf{y}_{N-2}\\
%   \vdots             & \ddots   & \vdots\\
%   \mathbf{y}_{1}     & \dots    & \mathbf{y}_{N-p}\\
%   1                  & \dots    & 1 
%   \end{bmatrix}^\top \, ,\\
%\mathbf{B} &=&
%\begin{bmatrix}
%\quad \\
%\mathbf{y}_{p+1} &
%\dots & 
%\mathbf{y}_N \\
%\quad
%\end{bmatrix}^\top  \, ,\\
%\mathbf{X}&=&
%\left[
%\begin{array}{ccccc}
%    \quad  & \quad & \quad & \quad \\
%    \phi_1   & \cdots & \phi_p & \mathbf{c} \\  
%    \quad & \quad & \quad & \quad
%    \end{array}
%\right]^\top \, ,\\
%\mathbf{E} &=& 
%\begin{bmatrix}
%    \quad \\
%    \mathbf{\epsilon}_{p+1}  & 
%    \dots                & 
%    \mathbf{\epsilon}_N \\
%    \quad
%\end{bmatrix}^\top \, . 
%\end{eqnarray}

\begin{equation}
 \label{eq:vareq}
               \underbrace{ \begin{bmatrix}
               \quad \\
               \mathbf{y}_{p+1} &
               \mathbf{y}_{p+2} &
               \dots & 
               \mathbf{y}_N \\
               \quad
               \end{bmatrix}}_{\substack{ \mathbf{B}^\top\\l \times (N-p)}}   
= 
                \underbrace{\left[ 
                \begin{array}{ccccc}
                \quad & \quad & \quad & \quad & \quad \\
                \phi_1  & \phi_2 & \cdots & \phi_p & \mathbf{c} \\  
                \quad &\quad & \quad & \quad & \quad
               \end{array} 
               \right]}_{\substack{ \mathbf{X}^\top\\ l \times (l \times p + 1 )}}
\underbrace{\begin{bmatrix}
   \mathbf{y}_{p}  & \mathbf{y}_{p+1} & \dots    & \mathbf{y}_{N-1}\\
   \mathbf{y}_{p-1}  & \mathbf{y}_{p} & \dots    & \mathbf{y}_{N-2}\\
   \vdots        & \vdots   & \ddots   & \vdots\\
   \mathbf{y}_{1} & \mathbf{y}_{2}   & \dots    & \mathbf{y}_{N-p}\\
   1 & 1   & \dots    & 1 
   \end{bmatrix}}_{\substack{ \mathbf{A}^\top\\ (l\times p +1 )\times (N-p)}}
+
\underbrace{\begin{bmatrix}
                \quad \\
              \mathbf{\epsilon}_{p+1}  & 
              \mathbf{\epsilon}_{p+2}  & 
              \dots                & 
              \mathbf{\epsilon}_N \\
              \quad
             \end{bmatrix}}_{\substack{\mathbf{E}^\top\\l \times (N-p) }} 
\end{equation}
Equation~(\ref{eq:vareq})  can be solved using ordinary least squares estimation.

\subsection{Vector Error Correction}

VECM is a special form of a VAR model for I(1) variables that are also
cointegrated~\cite{banerjee1993}. It is obtained by replacing the form $\Delta
\mathbf{y}_t = \mathbf{y}_t - \mathbf{y}_{t-1}$ in equation (\ref{eq:var}).
VECM is expressed in terms of differences, has an error correction term and 
the following form:

\begin{equation}
 \label{eq:vec}
 \Delta \mathbf{y}_t = 
 \underbrace{ \Omega\mathbf{y}_{t-1}}_\text{Error correction term} + 
 \sum_{i=1}^{p-1}
\phi_i^* \Delta \mathbf{y}_{t-i}  + \mathbf{c} + \mathbf{\epsilon}_t \quad ,
\end{equation}

\noindent where coefficients matrices $\Omega$ and $\phi_i^*$ are
function of matrices $\phi_i$ (shown in equation (\ref{eq:var})) as follows:

\begin{eqnarray*}
\phi_i^* &: =& -\sum_{j=i+1}^{p} \phi_j \\
\Omega &: =& -(\mathbb{I}-\phi_1-\dots-\phi_p) 
\end{eqnarray*}

The matrix $\Omega$ has the following properties~\cite{johansen1995}:
\begin{itemize}
\item If $\Omega = 0$ there is no cointegration
\item If $rank(\Omega)=l$ i.e full rank, then the time series are not
I(1) but stationary
\item If $rank(\Omega)=r,\quad 0 < r < l$ then, there is cointegration
and the matrix $\Omega$ can be expressed as $\Omega =
\alpha \beta^\top$, where $\alpha$ and $\beta$ are $(l \times r)$
matrices and $rank(\alpha)=rank(\beta)=r$.

The columns of $\beta$ contains the cointegration vectors and the rows of
$\alpha$ correspond with the adjusted vectors. $\beta$ is obtained by Johansen
procedure~\cite{johansen1988} whereas $\alpha$ has to be determined as a
variable in the VECM.

It is worth noticing that the factorization of the matrix $\Omega$ is not
unique since for any $r \times r$ nonsingular matrix $H$ we have:

\begin{eqnarray*}
\alpha \beta^\top &=& \alpha \mathbf{HH^{-1}} \beta^\top\\
&=&(\alpha\mathbf{H})(\beta(\mathbf{H}^{-1})^\top)^\top \\
&=& \alpha^*(\beta^*)^\top
\end{eqnarray*}

\noindent whith $\alpha^* = \alpha\mathbf{H}$ and $\beta^* =
\beta(\mathbf{H}^{-1})^\top$.

\end{itemize}

If cointegration exists, then equation (\ref{eq:vec}) can be written as follows:

\begin{equation}
 \label{eq:vecfull}
 \Delta \mathbf{y}_t = \alpha \beta^\top\mathbf{y}_{t-1} 
 + \sum_{i=1}^{p-1} \phi_i^*\Delta
\mathbf{y}_{t-i}  + \mathbf{c} + \mathbf{\epsilon}_t \quad ,
\end{equation}

\noindent which is a VAR model but for time series differences.

VECM matricial form is the following:

%\begin{eqnarray}\label{eq:vecmatrix}
% \mathbf{A}&=&
%   \begin{bmatrix} 
%   \beta^\top \mathbf{y}_{p} & 
%   \cdots & \beta^\top \mathbf{y}_{N-1} \\
%   \Delta \mathbf{y}_p  & \cdots 
%   &\Delta\mathbf{y}_{N-1} \\ 
%   \vdots  & \ddots & \vdots \\
%   \Delta\mathbf{y}_2  & \cdots 
%   & \Delta \mathbf{y}_{N-p+1} \\
%   1 & \cdots & 1 
%   \end{bmatrix}^\top \label{eq:vecA} \, ,\\
%\mathbf{B} & = &
% \begin{bmatrix}
% \quad\\
%  \Delta \mathbf{y}_{p+1} & 
%  \dots &
%  \Delta \mathbf{y}_N \label{eq:vecB}\\
%  \quad
% \end{bmatrix}^\top \, ,\\
%\mathbf{X}&=&
%  \begin{bmatrix}
%   \quad \\
%   \alpha & \phi_1^* & \cdots & \phi_{p-1}^* & \mathbf{c} \\  
%   \quad
%   \end{bmatrix}^\top \, ,\\
%\mathbf{E} &=&
%\begin{bmatrix}
%   \quad \\
%   \mathbf{\epsilon}_{p+1} &
%   \dots &
%   \quad &\mathbf{\epsilon}_N \, \\ \quad
%\end{bmatrix}^\top 
%\end{eqnarray}
%

\begin{equation} \label{eq:vecmatrix}
\underbrace{
               \begin{bmatrix}
               \quad\\
                \mathbf{\Delta y}_{p+1} & 
                \dots &
                \mathbf{\Delta y}_N \\
                \quad
               \end{bmatrix}
               }_{\substack{\mathbf{B}^\top\\ l \times (N-p) }} =
   \underbrace{
    \begin{bmatrix}
     \quad \\
     \alpha & \phi_1^* & \cdots & \phi_{p-1}^* & \mathbf{c} \\  
     \quad
     \end{bmatrix} 
     }_{\substack{ \mathbf{X}^\top\\ l \times (l\times p +1)}}
\underbrace{\begin{bmatrix} 
   \beta^\top \mathbf{y}_{p} & 
   \cdots & \beta^\top \mathbf{y}_{N-1} \\
   \mathbf{\Delta y}_p  & \cdots 
   & \mathbf{\Delta y}_{N-1} \\ 
   \vdots  & \ddots & \vdots \\
   \mathbf{\Delta y}_2  & \cdots 
   & \mathbf{\Delta y}_{N-p+1} \\
   1 & \cdots & 1 
   \end{bmatrix}}_{\substack{\mathbf{A}^\top \\ (l \times p +1) \times (N-p) }}
+
\underbrace{\begin{bmatrix}
              \quad \\
              \mathbf{\epsilon}_{p+1} &
              \dots &
              \quad &\mathbf{\epsilon}_N \\ \quad
             \end{bmatrix}}_{\substack{\mathbf{E}^\top\\ l \times (N-p) }} 
\end{equation}

VAR and VECM parameters shown in equations (\ref{eq:vareq},\ref{eq:vecmatrix}) can be solved
using standard regression techniques, such as ordinary least squares (OLS).

@@

When $\mathbf{A}$ is singular, solution to equation~(\ref{eq:vecmatrix}) is given
by the ordinary least squares (OLS) method. OLS consists of minimizing the sum
of squared errors or equivalently minimizing the following expression:

@@

\subsection{Ridge Regression}\label{sec:RR}

Ridge regression (RR) introduces a ridge estimation to avoid the
singularity of the matrix $\mathbf{A}^\top \mathbf{A}$
equation~(\ref{eq:pseudoinverse}). RR solves the following
optimization problem:

@@

\subsection{Selection of Lambda}

The additional term $\lambda \|\mathbf{\mathbf{X}}\|_2^2$ in the optimization
problem shown in equation~(\ref{eq:RRproblem}) has two effects on the solution:
shrinks the coefficients towards zero and improves the conditioning of the
problem.

On the other hand, the effect of adding the term $\lambda \mathbb{I}$
to the matrix $\mathbf{A}^\top \mathbf{A}$
(equation~(\ref{eq:optsolRR})) improves its condition number since it
increases its diagonal values when $\lambda > 0 $.  The matrix
$\mathbf{A}^\top \mathbf{A}$ is symmetrical ($(\mathbf{A}^\top
\mathbf{A})^\top = \mathbf{A}^\top \mathbf{A}$) and therefore
diagonalizable.  If we know the eigenvalue decomposition of $\mathbf{A}
= \mathbf{U\Sigma V^\top}$, then:

\begin{eqnarray*}
\mathbf{A}^\top \mathbf{A}+\lambda \mathbb{I}&=&\mathbf{V\Sigma^2
V}^\top + \lambda \mathbf{V} \mathbf{V}^\top\\ &=&\mathbf{V}
(\mathbf{\Sigma}^2+\lambda\mathbb{I}) \mathbf{V}^\top \, ,
\end{eqnarray*}

\begin{equation*}
\mathbf{\Sigma}^2+\lambda\mathbb{I}=
\begin{bmatrix}
\sigma^2_1 + \lambda & \, & \, \\
\, & \sigma^2_2 +\lambda & \, \\
\, & \, & \ddots & \, \\
\, & \, & \, & \sigma^2_n +\lambda \, .
\end{bmatrix}
\end{equation*}

where $\sigma_1 \geq \sigma_2 \geq \dots \geq \sigma_n$.

Since the condition number of a matrix $\mathbf{A}$ is defined as:

\begin{equation*}
	\kappa = \|\mathbf{A}\| \|\mathbf{A}^{-1}\|
\end{equation*}

If matrix $\mathbf{A}$ is non-singular, its condition number can be
expressed in terms of its singular values. The effect of adding the
regularization term affects the condition number as follows:

\begin{eqnarray*}
\kappa_{ols} &=& \|\mathbf{A}\| \|\mathbf{A}^{-1}\|=\frac{\sigma_1}{\sigma_n} \\
\kappa_{ridge} &=& \|\mathbf{A}^\intercal \mathbf{A} + \lambda \mathbb{I}\| 
\|(\mathbf{A}^\intercal \mathbf{A} + \lambda \mathbb{I})^{-1}\|=\frac{\sigma_1+\lambda}{\sigma_n + \lambda} \,
\end{eqnarray*}

It is easy to see that the term $\lambda$ improves the condition number:

\begin{equation*}
        \frac{\sigma_1+\lambda}{\sigma_n + \lambda} <
        \frac{\sigma_1}{\sigma_n} \,  \qquad \forall \quad \lambda > 0
\end{equation*}

However, $\lambda$ cannot be too large. Tipically $\lambda$ is small and its
magnitude depends on the matrix $\mathbf{A}$.

For rank deficient matrices we know that $\text{det}(\mathbf{A A^\top})=0$, adding the
term $\lambda \mathbb{I}$ we have that $\text{det}(\mathbf{A A^\top}+\lambda \mathbb{I}) =
p(\lambda)$ where $p(\lambda)$ is a polynomial of degree $n$ ($\mathbf{A}$ is
$m \times n$). The zeros of $p(\lambda)$ are discretes, so it can be represented
as:

\[
p(\lambda) =
\lambda(\lambda-\lambda_1)^{n_1}(\lambda-\lambda_2)^{n_2}\dots(\lambda-\lambda_s)^{n_s}
\]

\noindent where $n_1 + n_2 + \dots + n_s = n$.

This means that $\lambda$ must be small in order to ensure that $p(\lambda)$
does not vanish.

\subsubsection{The bias-variance tradeoff}

To determine parameter $\lambda$ is crucial for ridge regression since it could
reduce the expected prediction error by reducing variance considering a biased
estimator, this is called the bias-variance trafeoff. The prediction error is
obtained as:

\begin{equation}
 \label{eq:prederror}
E[(B-\hat{f}(\mathbf{X}))^2] = \sigma^2 + Bias(\hat{f}(\mathbf{X}))^2 + Var(\hat{f}(\mathbf{X}))
\end{equation}

\noindent where $\hat{f}(\mathbf{X})=\mathbf{AX}(\lambda)$ and
$f(\mathbf{X})=\mathbf{AX}$.

Ridge regression shows an
increasing bias and a decreasing variance. Despite OLS has zero bias,
its variance is greater than ridge for small values of $\lambda$.
It can be shown that, in terms of prediction error, ridge (black line)
is lower than OLS (dotted line)~\cite{hoerl1970}.

@@

Despite the bias-variance tradeoff provides a conceptual framework for
determining a good model, is not directly useful. Some popular methods
for determine model selection are:

@@

On the other hand, $\lambda$ can also be used to reduce computational time as it
is shown in equation~(\ref{eq:taylor}). Golub (~\cite{golub1965}) suggested that
$\lambda$ had to be chosen so that:

\[
\frac{\lambda}{\lambda + \delta^2} < 0.1
\]

\noindent where $\delta$ is a lower bound of the smallest non-zero singular
value $\sigma_k$. One approach is to choose the greatest lower bound for
$\delta=\sigma_k$. Hence:

\[
\frac{\lambda}{\lambda + \sigma_k^2} < 0.1
\]

\noindent if we choose $\lambda=\beta \sigma_k^2$ then we get the following
relation:

\[
\frac{\beta \sigma_k^2}{\beta \sigma_k^2 + \sigma_k^2} =
\frac{\beta}{\beta+1} < 0.1 \Rightarrow \beta < \frac{1}{9}  
\]

Experiments done by ~\cite{coleman+sun2010} have shown that $\beta = 0.01$
produces satisfactory results. This means that $\lambda = 0.01 \sigma_k^2$.

However, get $\sigma_k$ imply obtaining first the SVD, which is
computational expensive.

Since the algorithm presented by ~\cite{coleman+sun2010} already compute the QR
factorization of the matrix $\mathbf{A}$ they show a procedure for getting
a $\lambda$ approximation using QR.

\begin{enumerate}
\item Compute QR factorization: $\mathbf{A} = \mathbf{Q_1R_1}$
\item Let $\mathbf{W}$ denote the set of absolute values of the nonzero diagonal
elements of $\mathbf{R_1}$. Let $w_{\text{min}}$ and $w_{\text{max}}$ denote the
smallest and largest elements of $\mathbf{W}$ respectively. Both 
$\lambda_1 = \hat{\beta} w_{\text{min}}^2$
$\lambda_2 = \hat{\beta} \frac{w_{\text{min}}^2}{w_{\text{max}}^2}$ where
$\hat{\beta}=0.00025$ produce satisfactory results.
\item The $\lambda_{\text{QR}}$ approximation is obtained as:
$\lambda_{\text{QR}} = \frac{\lambda_1 + \lambda_2}{2}$  
\end{enumerate}

\subsection{Efficient computation}

%Revisar!!!
In regression problems we always require getting a matrix inverse which is
computationally expensive. However, in stream data problems the is a way to
obtain a matrix inverse approximation using the Sherman-Morrison-Woodbury. This
formula allows to get the inverse of a
the matrix $\mathbf{A+uv^\top}$ if we already calculated the
inverse of $\mathbf{A}$ as follows:

\begin{equation}
\label{eq:SMW}
(\mathbf{A+uv^\top})^{-1}=\mathbf{A}^{-1}-
\frac{\mathbf{A^{-1}uv^\top A^{-1}}}{\mathbf{1+v^\top A^{-1}u}}
\end{equation}

%Revisar!!

Another way was presented by Coleman and
Sun~\cite{coleman+sun2010} using an iterative algorithm which
uses $\mathbf{X}(\lambda)$ to approximate $\mathbf{X}(0)$.

Using the compact SVD (shown in equation~(\ref{eq:compactsvd}))
$\mathbf{X}(\lambda)$ is expressed as follows:

\begin{eqnarray}
\label{eq:optsolRRsvd}
\mathbf{X}(\lambda) & = & (\mathbf{A}^\top \mathbf{A}+ \lambda
\mathbb{I})^{-1}\mathbf{A}^\top \mathbf{B} \nonumber \\
& = &\mathbf{V}_1(\Sigma_1^2+\lambda \mathbb{I})^{-1}\mathbf{\Sigma_1
U_1^\top B}
\end{eqnarray}

\noindent where is easy to see that $\mathbf{X}(\lambda) \rightarrow
\mathbf{\hat{X}}=\mathbf{V}_1 \mathbf{\Sigma}_1^{-1}\mathbf{U}_1^\top
\mathbf{B}$ as $\lambda \rightarrow 0$.

The method consists in obtaining $\mathbf{X}(\lambda)$ and then refine
by adding more terms of its Taylor expansion to approximate
$\mathbf{X}(0) = \mathbf{\hat{X}}$. The Taylor expansion about $\lambda_0$ is:

\begin{equation}
\label{eq:taylor}
    \mathbf{X}(\lambda)=\mathbf{X}(\lambda_0) + \sum_{k=1}^\infty
    \mathbf{s}_k(\lambda-\lambda_0)^{k}
\end{equation}

\noindent where $\mathbf{s}_k=\frac{1}{k!}\mathbf{X}(\lambda)^{(k)}$
and $\mathbf{X}(\lambda)^{(k)}$ is obtained by taking differences 
$\frac{\partial}{\partial \lambda}$ of:

\begin{eqnarray*}
(\mathbf{A}^\top \mathbf{A}+ \lambda\mathbb{I}) \mathbf{X}(\lambda) & = & \mathbf{A}^\top \mathbf{B}\\
(\mathbf{A}^\top \mathbf{A}+ \lambda\mathbb{I}) \mathbf{X}(\lambda)^{(1)} + \mathbf{X}(\lambda)& = & 0 \\
\mathbf{X}(\lambda)^{(1)}  &=& -(\mathbf{A}^\top \mathbf{A}+ \lambda\mathbb{I}) ^{-1} \mathbf{X}(\lambda) \\
\mathbf{X}(\lambda)^{(2)}  &=& -2(\mathbf{A}^\top \mathbf{A}+ \lambda\mathbb{I}) ^{-1} \mathbf{X}(\lambda)^{(1)} \\
& \vdots & \\
\mathbf{X}(\lambda)^{(k)}  &=& -k(\mathbf{A}^\top \mathbf{A}+ \lambda\mathbb{I}) ^{-1} \mathbf{X}(\lambda)^{(k-1)} 
\end{eqnarray*}

\noindent for $\lambda=0$ we have:

\begin{equation}
\label{eq:taylor}
    \mathbf{X}(0)=\mathbf{X}(\lambda_0) + \sum_{k=1}^\infty
     (-1)^k \mathbf{s}_k \lambda_0^k
\end{equation}

\noindent therefore in order to ensure convergence, we can see that$\lambda_0$
selection cannot be large.

The algorithm for computing $\mathbf{X}(0)$ is the following:

\begin{algorithm}[H]
\begin{algorithmic}[1]
\REQUIRE $\,$ \\
$\mathbf{A}$: design matrix \\
$\mathbf{B}$: response matrix \\
$\lambda$: rank deficient parameter \\
\ENSURE  $\,$ \\
$\mathbf{X}$: parameters \\
\STATE $\mathbf{M}=\mathbf{A^\top A}$ \\
\STATE Initialize $\mathbf{Q R}=\mathbf{M}$ \\
\STATE $\mathbf{X} = \mathbf{R^{-1}Q^\top B}$ \\
\STATE $\mathbf{s} = \mathbf{X}$ \\
\FOR { $i = 1,2,3,\dots$ }
	\STATE $\mathbf{s} =
        -(\mathbf{M}+\lambda\mathbb{I})^{-1}\mathbf{s}$\\
	\STATE $\mathbf{X}=\mathbf{X} + (-1)^i {\color{red}\mathbf{s}
        \lambda^i}$
\ENDFOR
\end{algorithmic}
\caption{Algorithm for handling rank deficient matrices}
\label{alg:coleman}
\end{algorithm}

The algorithm~\ref{alg:coleman} solves equation~(\ref{eq:taylor}). However, this
version is unstable since tipically $\|\mathbf{s}\|$ is very large and
$\lambda^i$ is very small ($\lambda$ is small).

The following algorithm shows a more stable version of
algorithm~\ref{alg:coleman}.

\begin{algorithm}[H]
\begin{algorithmic}[1]
\REQUIRE $\,$ \\
$\mathbf{A}$: design matrix \\
$\mathbf{B}$: response matrix \\
$\lambda$: rank deficient parameter \\
\ENSURE  $\,$ \\
$\mathbf{X}$: parameters \\
\STATE $\mathbf{M}=\mathbf{A^\top A}$ \\
\STATE Initialize $\mathbf{Q R}=\mathbf{M}$ \\
\STATE $\mathbf{X} = \mathbf{R^{-1}Q^\top B}$ \\
\STATE $\mathbf{t} = \mathbf{X}$ \\
\FOR { $i = 1,2,3,\dots$ }
        \STATE $\mathbf{t} =\lambda \mathbf{t}$  
        \STATE $\mathbf{t} =  -(\mathbf{M}+\lambda\mathbb{I})^{-1}\mathbf{t}$
	\STATE $\mathbf{X}=\mathbf{X} + \mathbf{t}$
\ENDFOR
\end{algorithmic}
\caption{Algorithm for handling rank deficient matrices improved}
\label{alg:colemanimproved}
\end{algorithm}

Both algorithms are equivalent, but algorithm~\ref{alg:colemanimproved} is more
stable and converges in tipically less than 10 steps.
It is important to notice that the QR factorization is computed only once and it
is computationally less expensive than the SVD.

>>>>

VECM is a special case of the VAR \cite{sims1980} model and both describe the joint behaviour
of a set of variables.

The VAR($p$) model is a general framework describing the behaviour of a
set of $l$ endogenous variables as a linear combination of their last $p$
values, where $l,p\in\mathbb{N}$. 
In our case, each one of these $l$ variables is a scalar time series
$y_{\lambda,t}$, $\lambda=1,\dots,l$, and we represent them all together
at time $t$ by the the vector time series:
\begin{equation}
\label{eq:variables}
\mathbf{y}_t = 
\begin{bmatrix} y_{1,t} & y_{2,t} & \dots & y_{l,t} \end{bmatrix}^\top.
\end{equation}
\noindent
Notice that the vectors $\mathbf{y}_t$ are assumed to be $l$-dimensional.

The VAR($p$) model describes the behaviour of a dependent variable in terms of
its own lagged values and the lags of the others variables in the system. The
model with $p$ lags is formulated as the system of $N$:
\begin{align}
\label{eq:var}
\mathbf{y}_t 
= \boldsymbol{\Phi}_1 \mathbf{y}_{t-1} +
  \boldsymbol{\Phi}_2 \mathbf{y}_{t-2} + \dots +
  \boldsymbol{\Phi}_p\mathbf{y}_{t-p} +
  \mathbf{c} + \boldsymbol{\epsilon}_t \nonumber \\
t=p+1,\dots,N,
\end{align}
\noindent where 
$\boldsymbol{\Phi}_1, \boldsymbol{\Phi}_2,\dots,\boldsymbol{\Phi}_p$
are $l\times l$-matrices of real coefficients,
$\boldsymbol{\epsilon}_{p+1},
 \boldsymbol{\epsilon}_{p+2}, \dots, \boldsymbol{\epsilon}_N$ 
are error terms, $\mathbf{c}$ is a constant vector and $N$ is the total
number of samples.

Notice that, regarding our notation of section (\ref{sec:coint}),
we have here 
$\mathbf{y}_t^0 = \mathbf{y}_t$,
$\mathbf{y}_t^\nu = \mathbf{y}_{t-\nu}$ and
the $\lambda$-th component of the vector time series $\mathbf{y}_t^\nu$
is the scalar time series $y_{\lambda,t}^\nu$, where $\nu=1,\dots,p$ and
$\lambda=1,\dots,l$.

However, the VAR model cannot be used with non-stationary variables. VECM \cite{engle87} is also a linear model and a special form of a VAR model for I(1) variables that are also
cointegrated~\cite{banerjee1993}. If cointegration exists, variable differences
are stationary and they introduce an error correction term which adjusts
coefficients to bring the variables back to equilibrium.

It is obtained re-writing equation (\ref{eq:var}) in terms of the new
variable $\Delta\mathbf{y}_t=\mathbf{y}_t-\mathbf{y}_{t-1}$.
The VECM model, expressed in terms those differences, takes the form:
\begin{equation}\label{eq:vec}
\Delta \mathbf{y}_t 
= \boldsymbol{\Omega}\,\mathbf{y}_{t-1}
  + \sum_{i=1}^{p-1} \boldsymbol{\Phi}_i^*\,\Delta\mathbf{y}_{t-i}
  + \mathbf{c} + \boldsymbol{\epsilon}_t\,,
\end{equation}
\noindent
where the coefficients matrices $\boldsymbol{\Phi}_i^*$ and 
$\boldsymbol{\Omega}$, expressed in terms of the matrices
$\boldsymbol{\Phi}_i$ of (\ref{eq:var}), are:
\begin{align*}
\boldsymbol{\Phi}_i^* 
&:= -\sum_{j=i+1}^{p}\boldsymbol{\Phi}_j\,, \\
\boldsymbol{\Omega}
&:= -\left( \mathbb{I} - \boldsymbol{\Phi}_1 - \dots 
    - \boldsymbol{\phi}_p \right)\,. 
\end{align*}
The following well known properties of the matrix $\boldsymbol{\Omega}$
\cite{johansen1995} will be useful in the sequel:
\begin{itemize}
\item
If $\boldsymbol{\Omega} = \mathbf{0}$, there is no cointegration.
\item 
If $rank(\boldsymbol{\Omega})=l$, i.e., if $\boldsymbol{\Omega}$ has
full rank, then the time series are not I(1) but stationary.
\item
If $rank(\boldsymbol{\Omega})=r$, $0<r<l$, then there is cointegration
and the matrix $\boldsymbol{\Omega}$ can be expressed as
$\boldsymbol{\Omega}=\boldsymbol{\alpha\beta}^\top$, where $\boldsymbol{\alpha}$
and $\boldsymbol{\beta}$ are
$l\times r$ matrices and
$\text{rank}(\boldsymbol{\alpha})=\text{rank}(\boldsymbol{\beta})=r$.
\item
The columns of $\boldsymbol{\beta}$ contains the cointegration vectors and the rows of
$\boldsymbol{\alpha}$ correspond with the adjusted vectors. 
$\boldsymbol{\beta}$ is obtained by Johansen procedure~\cite{johansen1988},
whereas $\boldsymbol{\alpha}$ has to be determined as a variable in the VECM.
\end{itemize}
It is worth noticing that the factorization of the matrix
$\boldsymbol\Omega$ is not unique, since for any $r \times r$
non-singular matrix $\mathbf{H}$, $\boldsymbol{\alpha}^*:=\boldsymbol{\alpha}\mathbf{H}$,
and $\boldsymbol{\beta}^*=\boldsymbol{\beta}(\mathbf{H}^{-1})^\top$ we have
$\boldsymbol{\alpha\beta}^\top=\boldsymbol{\alpha}^*(\boldsymbol{\beta}^*)^\top$.
If cointegration exists, then equation (\ref{eq:vec}) can be written
as follows:
\begin{equation}\label{eq:vecfull}
\Delta\mathbf{y}_t 
= \boldsymbol{\alpha\beta}^\top\mathbf{y}_{t-1} 
  + \sum_{i=1}^{p-1}\boldsymbol{\Phi}_i^*\,\Delta\mathbf{y}_{t-i}
  + \mathbf{c} + \boldsymbol{\epsilon}_t\,,
\end{equation}
\noindent
which is a VAR model but for time series differences.

Transposing each equation of the system (\ref{eq:vecfull}) we can write
the VECM($p$) model in block-matrix form as:
\begin{equation}\label{eq:vareq}
\mathbf{B} = 
\mathbf{A} \mathbf{X} + 
\mathbf{E} \, , 
\end{equation}
%
\noindent where $\mathbf{B}$ dimension is $((N-p)\times l)$, $\mathbf{A}$
dimension is $((N-p)\times(r+(p-1)l +1))$, $\mathbf{X}$ dimension is $((r+(p-1)l
+1)\times l)$ and $\mathbf{E}$ dimension is $((N-p)\times l)$:
%
\begin{alignat}{3}
\mathbf{B}
&= \begin{bmatrix}
   \Delta\mathbf{y}_{p+1}^\top \\
   \Delta\mathbf{y}_{p+2}^\top \\
   \vdots \\
   \Delta\mathbf{y}_N^\top
   \end{bmatrix}
&\quad
\mathbf{X}
&= \begin{bmatrix}
   \boldsymbol{\alpha}^\top \\
   \boldsymbol{\Phi}_1^{*\top} \\
   \boldsymbol{\Phi}_2^{*\top} \\
   \vdots \\
   \boldsymbol{\Phi}_{p-1}^{*\top} \\
   \mathbf{c}^\top
   \end{bmatrix}
&\quad
\mathbf{E}
&= \begin{bmatrix}
   \boldsymbol{\epsilon}_{p+1}^\top \\
   \boldsymbol{\epsilon}_{p+2}^\top \\
   \vdots \\
   \boldsymbol{\epsilon}_N^\top \\
   \end{bmatrix}
\end{alignat}
\noindent and 
\begin{align}
\mathbf{A} 
&= \begin{pmat}[{....|}]
   \mathbf{y}_p^\top \boldsymbol{\beta} & \Delta \mathbf{y}_p^\top & \Delta\mathbf{y}_{p-1}^\top & \dots 
                    & \Delta\mathbf{y}_2^\top & 1 \cr
   \mathbf{y}_{p+1}^\top  \boldsymbol{\beta} &\Delta\mathbf{y}_{p+1}^\top & \Delta\mathbf{y}_p^\top & \dots
                       & \Delta\mathbf{y}_3^\top & 1 \cr
   \vdots & \vdots & \vdots & \ddots & \vdots & \vdots \cr
   \mathbf{y}_{N-1}^\top  \boldsymbol{\beta} &\Delta\mathbf{y}_{N-1}^\top & \Delta\mathbf{y}_{N-2}^\top & \dots 
                       & \Delta\mathbf{y}_{N-p-1}^\top & 1 \cr
   \end{pmat}\, .
\label{eq:Amatrix}
\end{align}
Taking into account the error term $\mathbf{E}$, equation~(\ref{eq:vareq}) 
can be solved with respect to $\mathbf{X}$ using the ordinary least
squares (OLS) estimation.

@@

When $\mathbf{A}$ is singular, solution to equation~(\ref{eq:vareq}) is given
by the ordinary least squares (OLS) method. OLS consists of minimizing the sum
of squared errors or equivalently minimizing the following expression:

@@

\subsection{Regularization}\label{sec:RR}

Regularization can be understood using two contexts: learning and inverse problems. In the context of learning, regularization refers to techniques allowing to avoid over-fitting, e.g. ridge regression (RR) or Tikhonov regularization \cite{ tikhonov1977}

introduces a ridge estimation to avoid the
singularity of the matrix $\mathbf{A}^\top \mathbf{A}$
equation~(\ref{eq:pseudoinverse}). RR solves the following
optimization problem:

which was independently by Tikhonov \cite{tikhonov1963} and Phillips \cite{phillips1962}.

@@

\subsection{Regularization parameter chose}

The additional term $\lambda \|\mathbf{\mathbf{X}}\|_2^2$ in the optimization
problem shown in equation~(\ref{eq:RRproblem})
improves the conditioning of the problem because is equivalent to add the term $\lambda \mathbb{I}$ to the matrix $\mathbf{A}^\top \mathbf{A}$
(equation~(\ref{eq:optsolRR})) which effect is to improve its condition number.

To determine parameter $\lambda$ is crucial for ridge regression since it could
reduce the expected prediction error by reducing variance considering a biased
estimator, this is called the bias-variance tradeoff. 
For example, OLS has zero bias but its variance is greater than ridge regression for small values of $\lambda$. Additionally, in terms of prediction error, ridge regression is lower than OLS \cite{hoerl1970}.

@@

Despite the bias-variance tradeoff provides a conceptual framework for
determining lambda is not directly useful.

Some popular methods
for determine model selection are:

@@

On the other hand, $\lambda$ can also be used to reduce computational time as it
is shown in equation~(\ref{eq:taylor}). Golub (~\cite{golub1965})
